Details matter: Similar host galaxies can harbor different merging compact binary populations

### 2016-07-11 paper/paper.tex
Draft

@@

We parameterize the response function ${\cal G}(t|Z)$ with two factors: (a) an overall efficiency $\lambda$ for forming
binaries;  and (b) a
cumulative delay time distribution $P_t(<t|Z)$, describing the fraction of binaries born at $t=0$ that merge in less
than time $t$ from a population of binaries $Z$:\footnote{For simplicity, we assume the mass and delay time
  distributions are uncorrelated.  Figure A9 in \abbrvPSgrb{} shows this approximation, while not strictly true, is an
  excellent approximation for merging BH-BH binaries at solar metallicity.  }
\begin{eqnarray}
G(t,x|Z) &=& \lambda(Z) p_x(x|Z) \frac{dP_t}{dt}(t) 
\end{eqnarray}
Previous studies suggest the number of merging compact binaries $\lambda \delta M_*$ produced by a starburst depends
as a power law on metallicity:
\begin{eqnarray}
\label{eq:LambdaVersusZModel}
\lambda(Z) &=& \lambda_o \text{max}[(Z/Z_\odot)^{-a}, F_{max}] 
\end{eqnarray}
with $a\in [0,3]$ and $F_{max}<10^3$
%$\lambda_o F_{max} < 3\times 10^{-4}/M_\odot$
 (see, e.g.\cite{popsyn-LIGO-SFR-2008}, \abbrvKBLowZa).
In this paper we only investigate the dependence of rates with metallicity; we therefore  adopt 
\[
\lambda_0 = 10^{-3}/M_\odot
\]
as a typical value for neutron star compact binaries (see \abbrvPSgrb{}, \abbrvPSellipticals, and \abbrvKBLowZa).
%
Finally, for simplicity we adopt the universal delay time distribution
%% Finally, we parameterize the delay time distribution by  either
%% \begin{eqnarray}
%% P_t(<t) &=& \begin{cases}
%% 0 & t<10 \unit{Myr} \\
%% (t/13\unit{Gyr})^{b}& t \in [10\unit{Myr},t_H]
%% \end{cases} 
%% \end{eqnarray}
%% with $b> 0$ or as the special case
\begin{eqnarray}
\frac{dP_t(<t)}{dt} =  \begin{cases}
0 & t<10 \unit{Myr} \\
\frac{1}{t \ln (13 \unit{Gyr}/10\unit{Myr})} & t \in [10\unit{Myr},13\unit{Gyr}]
\end{cases} 
\end{eqnarray}
\abbrvPSgrb{} and \abbrvPSellipticals{} show this distribution is reasonable approximations to compact binary delay
time distributions.\footnote{Simulations suggest the delay time distribution varies from model to model and with mass.
  These variations have less impact on our results than the evolving metallicity distribution of star forming gas.}
%
For suitable choices of parameter, our phenomenological response function is qualitatively consistent with the results of detailed simulations of binary
evolution  \citep{2010ApJ...715L.138B,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,popsyn-LowMetallicityImpact2-StarTrackRevised-2012}.

\subsection{Tidal disruption and black hole mass versus metallicity}
At merger, black holes may promptly engulf a companion neutron star \editremark{same papers as for Brandon:Foucart,
  etc}, unless they have sufficiently small mass and sufficiently large spin.  Without  a tidal disruption event prior to merger, the prospects for associated electromagnetic counterparts
are severely diminished.  
%
A neutron star  (NS) coalescing with a black hole (BH) will be tidally disrupted prior to merger if and only if the component
masses, spins, and equation of state permit it.  As described below, only a small fraction of all  possible coalescing
BH-NS binaries have parameters consistent with tidal disruption.  
%
\jillianremark{Grammatical nightmare ahead: }Stellar evolution and hence supernovae remnants are expected to evolve
substantially with birth metallicity, the black hole mass and possibly spin in black hole-neutron star binaries can depend sensitively on binary formation
channels and the fraction of coalescing NS-BH binaries which are tidally disrupted will vary
over cosmic time. 
%
To quantify this effect, we adopt a simple formula for tidal disruption of neutron stars in BH-NS binaries and a simple
phenomenological model for the metallicity-dependent mass distribution of BHs in BH-NS binaries

\begin{figure}
\includegraphics{Figures/fig-mma-TidalDisrupt-MbhA-VersusTheta}
\caption{\label{fig:TidalDisrupt}\textbf{Tidal disruption}:  Following  \citet{2012PhRvD..86l4007F}, a plot of
  the critical spin needed for tidal disruption versus black hole mass and inclination (solid black, dotted black,
  dashed black, solid blue, and dotted blue are $\theta_{LS}=\pi/2,3\pi/8,\pi/4, \pi/8,0$ respectively).  Specifically,
  this figure corresponds to $M_{\rm disk}=0$ in Eq. (\ref{eq:Mdisk}), assuming  a $1.4 M_\odot$ neutron star with compactness ${\cal
    C}=1.66$ in inclined orbit around a spinning black hole of mass $m_{bh}$ and dimensionless spin $\chi_{bh}$.
  Parameters consistent with tidal disruption are above the curves; prompt capture occurs below the
  curve.
}
\end{figure}

% POINT: Trend versus mass and opening angle.  [Be careful: disruption time will depend on opening angle]
To estimate whether a specific compact binary will be tidally disrupted prior to merger, we employ the
following procecure to identify posterior samples consistent with tidal disruption of a neutron star.  First, the
smaller mass must lie within the range of masses allowed by our fiducial NS equation of state (\textbf{MPA1 from Ben}), here
between $0.5 M_\odot$ and $2.5 M_\odot$.  Second, the disruption process must leave behind a remnant disk with nonzero
mass, as estimated using the expression in \cite{2012PhRvD..86l4007F} (his Eqs. (6,12-13)):
\begin{align}
\frac{M_{disk}}{M_{NS}} &=
 0.288 (3 m_{bh}/m_{ns})^{1/3}[1-2 {\cal C}] 
\nonumber \\
&
 -0.148 (m_{bh}/m_{ns})  {\cal C} R_{isco}(a)/m_{bh} 
\label{eq:Mdisk}
\end{align}
where $q=m_{bh}/m_{NS}$ is the binary mass ratio; $a=S_{bh}/m_{bh}^2$ is a dimensionless measure of the black hole spin;
and  where ${\cal C}(m_{ns})=m_{ns}/R_{ns}$ is a mass- and equation-of-state dependent  measure of the neutron star
compactness.  In Foucart's expression, shown above, $R_{isco}(a)$ is the radius of the  innermost stable equitorial circular orbit  of a test particle
about a Kerr black hole  \citep{1972ApJ...178..347B}. % Brandon:  [their Eq. 2.21].   
When the black hole spin is not aligned with the orbit, we use the aligned component of the black hole spin magnitude $a\rightarrow a \cos \theta_{LS}$ in this
expression.   Fixing the neutron star mass $m_{ns}=1.4 M_\odot$ and hence compactness to ${\cal C}=1.66$, this expression implies a two-dimensional surface in
the three-dimensional parameter space $m_{bh},a,\theta_{LS}$ seperates parameters corresponding to tidal disruption to
parameters corresponding to prompt coalescence.
%
After coalescence, the remnant disk generates a tightly beamed electromagnetic counterpart.  For simplicity, we assume
beaming angle is independent of the disruption process or progenitor parameters, so a constant beaming correction factor $f_b$ enters into all
subsequent calculations and can hence be omitted.  
%
Figure \ref{fig:TidalDisrupt} shows contours of this expression for different values of $\theta_{LS}$.  \jillianremark{Add a sentence here describing the significance of the plot.  e.g. low-mass BHs can have a range of spins, regardless of angle, but high mass BHs always get high spins yay.}

\begin{figure}
\includegraphics{Figures/fig-mma-PrecomputeDisrupt}
\caption{\label{fig:TidalDisrupt:Probability}\textbf{Disruption fraction}: Contour plot of the fraction
  $P_{dis}(m_{bh},\chi_{bh})$ of BH-NS binaries with BH mass and spin $m_{bh},\chi_{bh}$ that are expected to tidally
  disrupt using the assumptions in the text.  Contours are shown at $0.1, 1/3, 0.5, 0.9$.}
\end{figure}

% POINT: Distribution of tilt angles
Based on Monte Carlo simulations of supernova kicks as described in the Appendix, we approximate the
cumulative tilt angle distribution $P(<\theta)$ of bound,
coalescing BH-NS binaries by the positive half of a normal distribution
\begin{eqnarray}
P(<\theta)\simeq \int_0^\theta e^{-\theta^2/2\sigma_\theta^2} \frac{2}{\sqrt{2\pi \sigma_\theta^2}} d\theta
\end{eqnarray}
where $\sigma_{\theta}\simeq 1$ is a characteristic angular width that varies relatively weakly with the black hole mass,
 supernova kick velocity, and compact binary formation channel.  This simple distribution is consistent with some detailed population synthesis simulations by
\texttt{StarTrack}.
For simplicity, in this work we adopt $\sigma_{\theta}=1$.   
%  POINT Average over spin distributions
Using this approximate distribution of misalignment angles, Figure \ref{fig:TidalDisrupt:Probability} estimates the fraction
$P_{dis}(m_{bh},\chi_{bh})$ of coalescing binaries with a given black hole mass and spin that should tidally disrupt.  
For simplicity assuming detection weighting does not strongly favor specific relative orientations $\theta$, this
distribution determines the relative likelihood that a gravitational wave event with a given BH mass and spin produces
is consistent with an electromagnetic counterpart.

% POINT: Average over distributions
The relative event rate of electromagnetic counterparts to BH-NS disruptions therefore depends on the intrinsic
distribution $p_d(m_{bh},\chi_{bh})$ of black hole mass and spin among GW-detected BH-NS mergers and the probability
that an event is detected via electromagnetic emission and GW: 
\begin{eqnarray}
p_{em}(m_{bh},\chi_{bh}) = \frac{P_{dis}(m_{bh},\chi_{bh}) p_d(m_{bh},\chi_{bh})}{\int dm_{bh} d\chi_{bh}P_{dis}(m_{bh},\chi_{bh}) p_d(m_{bh},\chi_{bh})}
\end{eqnarray}
where the constant beaming factor cancels.  
%
Because black holes in BH-NS binaries are currently expected to accrete relatively little mass, the spin distribution at merger
should be close to their birth spin distribution.   This distribution is entirely unknown but likely significant, given
tidally-induced spinup prior to merger\editremark{Ott}; we will assume the spin distribution to be uniform between $0.7$
and $1$, independent of formation channel or birth conditions: \editremark{change: make this one Z dependent}
\begin{eqnarray}
p(\chi_{bh}) = \frac{1}{0.3} \theta(\chi_{bh}-0.7)\theta(1-\chi_{bh})
\end{eqnarray}
As discussed in \cite{popsyn-LowMetallicityImpact2-StarTrackRevised-2012},  the black hole mass distribution depends
sensitively on the birth environment and coalescing binary formation channels; see their Tables 7 and 9 and their Figures 6,9, 11, and 13,
reconstructing the BH mass from the reported chirp masses (red curves).  
Omitting formation channels that predict implausibly high compact binary formation rates (their model A), the
distribution of $m_{bh}$ in BH-NS binaries extends over a wide range.  As a first approximation, the BH mass
distribution in BH-NS binaries does not depend on composition, favoring BH-NS binaries with relatively low BH masses
$m_{bh}\lesssim 5-9 M_\odot$ independent of birth composition.  
% \cite{2012ApJ...749...91F}  : physics of SN explosions used in these mechanisms. Rapid engine needed for mass gap, but
% is there one? 
% mc[m1_, m2_] := (m1 m2)^(3/5)/(m1 + m2)^(1/5)
% Solve[mc[m1, 1.4] == 2.5]
For simplicity, we will adopt a uniform mass distribution between $5$ and $9 M_\odot$:
\begin{eqnarray}
p(m_{bh}) = \frac{1}{4 M_\odot} \theta(m_{bh}-4 M_\odot)\theta(9 M_\odot-m_{bh})
\end{eqnarray}

% POINT: Include delay time: get examples of disruption fraction versus time

\section{Results}
\label{sec:results}

\begin{verbatim}
ACTION ITEMS AND STATUS

* Result script looks good (\texttt{sfr\_to\_rate.py}).  Follow instructions

** Modify script to include tidal disruption options for BHs

** plot 'stanard' results for NS-NS

** set scale to be physical units -- some ad-hoc factors used

*

* Do we get to different answers?  Are they different ENOUGH?

** Overall event rate: different by 20\% between two types. Largely similar.
   - plot versus z_exponent.  Is there a trend

** mass-cut-dependent event rate -- will be more sensitive (e.g., if there
  is a hard cutoff on some masses, as having no EM counterpart)

* Aside: BBH distribution the most sensitive, but not relevant for EM followup
\end{verbatim}

\appendix
\section{Binary evolution and spin-orbit misalignment}

\editremark{ROS to do} -- find python/C code (Davide?) for simple MC; mine old resuts from startrack

%% POINT: Other reference
% 2014arXiv1401.3032W  - fallback details

The progenitors of both components of a BH-NS binary undergo supernova, whose associated asymmetric mass loss and
supernova kicks significantly perturb the orbit  \citep{2013PhRvD..87j4028G,2002MNRAS.329..897H,2000ApJ...541..319K}.   Assuming progeniror stellar spins are
aligned with the orbital plane (e.g., due to strong tides), only asymmetric supernova kicks can change
the orbital plane and introduce misalignment between the BH spin and the orbit.   The distribution of spin-orbit
misalignments expected in any specific binary depends strongly on weakly-constrained binary evolution physics, the
treatment of which is beyond the scope of this article.  Previous detailed calculations have reported on spin-orbit
misalignment \citep{2010ApJ...719L..79F,2008ApJ...682..474B,2004PhRvD..69j2002G}.  That said, relatively simple order-of-magnitude calculations
can explain the typical small misalignment angles expected among merging BH-NS binaries.
% POINT: Small angle tilt forma (follows from simple considerations of angular momentum)
Specifically, in the limit of a small supernova kick $\vec{v}_K$, the orbital plane  change by an angle $\theta \simeq \vec{v}_K \cdot
\hat{z}/v_{orb}$ where $\hat{z}$ is the direction perpendicular to the (circular) orbital plane.  The black hole forms first and cannot
significantly change its spin magnitude or direction through accretion; therefore the two supernovae kicks produced by
forming the black hole and then neutron star add (in quadrature) to produce a characteristic RMS misalignment error
$\E{\theta^2}$.
%
For simplicity, we will adopt the model used in the \texttt{StarTrack} binary evolution code \citep{StarTrack2}: binaries
prior to each supernovae can be approximated as in circular orbits;  supernova kicks on neutron stars produce a Maxwellian
distribution of velocities with 1d velocity dispersion $\sigma_v=265$ \citep{HobbsKicks};  supernova kicks on black holes
$v_{bh}$ are suppressed by fallback, with velocities  $\vec{v}_{K,BH} = (1-f_b)v_{NS}$ suppressed by fallback accretion
onto the black hole; and the dominant formation channels for BH-NS binaries involve common
envelope accretion after the first supernova forms a black hole.   Combining these expressions under these assumptions, we expect the net misalignment  $\theta$ due to both supernova
explosions will have zero mean and variance
\begin{eqnarray}
\label{eq:MisalignmentOrderOfMagnitude}
\E{\theta^2} = \frac{\E{v^2}}{3}\left[ \frac{(1-f_b)^2\epsilon_1}{v_{orb,1}^2}+\frac{\epsilon_2}{v_{orb,2}^2} \right]
\end{eqnarray}
where $v_{orb,1}$ and $v_{orb,2}$ are the orbital velocities at the first and second supernova and where
$\epsilon_1,\epsilon_2$ are factors between 0 and 1 which account for the fact that the average is taken over binaries which \emph{survive} and are not
disrupted by the kick.  Because the second supernova  occurs in a tight orbit (post common
envelope),  the orbital velocity $v_{orb,2}$ at the supernova producing the neutron
star will be  large:  if $M_2$ and $r_2$ are the total mass and orbital separation, $v_{orb,2}=\sqrt{GM_{2}/r}\simeq
440\unit{km/s} \sqrt{M/M_\odot/(r_2/R_\odot)}$, 
% Convert[Sqrt[  GravitationalConstant (mbh SolarMass)/(r SolarRadius)], (Kilo Meter/    Second)]
comparable to or
significantly in excess of the typical supernova kick velocity for massive black holes orbiting the stripped progenitors of
neutron stars at seperations of a few $R_\odot$.  Conversely, at birth the black hole will often capture a significant
amount of the stellar envelope (``fallback'' accretion or prompt collapse) with $f_b\simeq 1$.  Therefore, even though
the initial orbit is wide, orbital  velocity is very small, and hence the orbit can be easily strongly misaligned by
supernova kicks,  the first kick when acting on a massive black hole can produce little to  modest amounts orbital
misalignment.

% POINT: Second term of order unity
For exceptional models of binary evolution, the population of coalescing BH-NS binaries will depend extremely critically on the
a finely-tuned second supernova.  In this case, only a small fraction of surviving BH-NS binaries merge; the BH-NS merger rate will be extremely low and  depend extremely sensitively on
the magnitude of supernova kicks; and the misalignment produced by the second supernova can frequently be very large.  More typically, however, a significant fraction of surviving BH-NS binaries merge
\citep{2008ApJ...675..566O,popsyn-LowMetallicityImpact2-StarTrackRevised-2012,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}; in these cases, the  second supernova's kick does not require fine-tuning to guarantee
coalescence.   In these most common scenarios, the second term in Eq. (\ref{eq:MisalignmentOrderOfMagnitude}) is of
order unity.

% POINT:
As a result, the process of surviving two supernova kicks generally produces a wide  distribution of misalignment
angles concentrated near $\theta\simeq 0$.  This wide distribution is expected to be relatively robust.  Extremely wide
distributions are disfavored kinematically and by the requirement that sufficiently many binaries survive to coalesce.
Conversely, the width of the distribution is bounded below by the contribution from the second supernova kick.\footnote{For simplicity and brevity, we neglect the
  possibility of different modes of supernova explosions and hence kicks, notably via electron capture supernovae \citep{2012ARNPS..62..407J,2004ApJ...612.1044P,2010MNRAS.406..656K,2015ApJ...801...32A}.  That said, small supernova kicks in NS
  birth would further minimize any spin-orbit misaligment and hence further favor the strong dependence of tidal disruption probability on the black hole mass and spin
  distribution described in the text.} 
As a first approximation,  the cumulative distribution of misalignment angles should agree with direct calculations \cite{2010ApJ...719L..79F,2008ApJ...682..474B,2004PhRvD..69j2002G}.

% POINT: 
%% To calculate the necessary inputs to this expression, we adopt one particular family of supernova mechanism:  Fryer et al 
%% \cite{2012ApJ...749...91F}, the model adopted in \texttt{StarTrack} and applied in detailed calculations of binary
%% evolution
%% \cite{popsyn-LowMetallicityImpact2-StarTrackRevised-2012,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}.
%% Because of significant residual uncertainty in the supernova mechanism, our results are proof-of-concept calculations
%% intended to illustrate plausible effects of composition on remnant masses, spins, and hence misalignments.  Other models
%% for supernovae produce dramatically different initial-final mass relations \cite{2012ApJ...757...69U} and are
%% anticipated to produce equally different different amounts of spin-orbit
%% misalignment in BH-NS binaries.  
%% %
%% Observations of isolated pulsar
%% recoils \cite{1997MNRAS.291..569H,Ar} and binary pulsars \cite{psr-kicks-WongWillemsKalogera2010}
%% strongly constrain the action of random supernova kicks on neutron stars.

\section{Black hole spins in BH-NS binaries}

* References: review in Fragos and McClintock \cite{2015ApJ...800...17F}  and references therein (LMXBs) -- low birth
spins, because of stripped companions; for more on measurement, see \cite{2011CQGra..28k4009M}

* References on BH spin versus mass: angular momentum capture in prompt collapse: O'Connor and Ott and references
therein (note no fallback); note Ugliano has a differen result. See also Clausen et
al (probability of BH vs mass) for more recent
%O'Connor and Ott: a criteria for explosions http://arxiv.org/pdf/1010.5550v2.pdf...note they don't have fallback.
%Clausen et al: trying to find formation probability vs mass http://arxiv.org/pdf/1406.4869v2.pdf (assuming no binaries)
%...note also http://adsabs.harvard.edu/abs/2012ApJ...757...69U (which is complicated)

%* References: \cite{2013ApJ...762..104S}

\section{Outline }
\begin{verbatim}

Intro
  -Important (edo/Brian)
  - More important soon (LIGO; IFU's give detail)
    because these galaxies are local and spatially resolved
    at 400 Mpc distance
  - The question: can we learn about things from detailed host galaxy 
    observations, and if so what?
    Obviously yes (cf. SN Ia, short GRB, etc)
  - Specific questions of interest, particularly for BH-NS origin for short GRBs:
    is there dependence of the event rate and BH mass on the source metallicity

  - Not clear because of kicks, mixing
  - Goal of paper: address this challenge via simulations

Methods
  - JB: specifically target two 'identical' MW galaxies (and their subhalo population)
  - Postproces JB: rhodot; Z of star forming gas
  - Model:
      - predict present-day number (and mass distribution) by formula

Results:
  - Different histories which look the same now can produce different answers
     - key issue: impact of past mergers
  -

Stuff to skip
  - (spin-dependent) tidal disruption threshold: assume all BH-NS disrupt, for simplicity
\end{verbatim}
%1 Kilo Parsec/(400 Mega Parsec)
%Convert[%, Milli ArcSecond] // N

\section{Other discussion}

@@

>>>>

%% absolutely critical vs optional
%
% critical
%  -write paper as is, minmal
% optional
%  - add dwarfs. They are important
%  - add ems counterparts. They are important
%  - discussion of multimessenger astro

@@

We parameterize the response function ${\cal G}(t|Z)$ with two factors: (a) an overall efficiency $\lambda$ for forming
binaries;  and (b) a
cumulative delay time distribution $P_t(<t|Z)$, describing the fraction of binaries born at $t=0$ that merge in less
than time $t$ from a population of binaries $Z$:\footnote{For simplicity, we assume the mass and delay time
  distributions are uncorrelated.  Figure A9 in \abbrvPSgrb{} shows this approximation, while not strictly true, is an
  excellent approximation for merging BH-BH binaries at solar metallicity.  }
\begin{eqnarray}
G(t,x|Z) &=& \lambda(Z) p_x(x|Z) \frac{dP_t}{dt}(t) 
\end{eqnarray}
Previous studies suggest the number of merging compact binaries $\lambda \delta M_*$ produced by a starburst depends
as a power law on metallicity:
\begin{eqnarray}
\label{eq:LambdaVersusZModel}
\lambda(Z) &=& \lambda_o \text{max}[(Z/Z_\odot)^{-a}, F_{max}] 
\end{eqnarray}
with $a\in [0,3]$ and $F_{max}<10^3$
%$\lambda_o F_{max} < 3\times 10^{-4}/M_\odot$
 (see, e.g.\cite{popsyn-LIGO-SFR-2008}, \abbrvKBLowZa).
In this paper we only investigate the dependence of rates with metallicity; we therefore  adopt 
\[
\lambda_0 = 10^{-3}/M_\odot
\]
as a typical value for neutron star compact binaries (see \abbrvPSgrb{}, \abbrvPSellipticals, and \abbrvKBLowZa).
%

For binaries containing neutron stars, for simplicity we adopt the universal delay time distribution
%% Finally, we parameterize the delay time distribution by  either
%% \begin{eqnarray}
%% P_t(<t) &=& \begin{cases}
%% 0 & t<10 \unit{Myr} \\
%% (t/13\unit{Gyr})^{b}& t \in [10\unit{Myr},t_H]
%% \end{cases} 
%% \end{eqnarray}
%% with $b> 0$ or as the special case
\begin{eqnarray}
\frac{dP_t(<t)}{dt} =  \begin{cases}
0 & t<10 \unit{Myr} \\
\frac{1}{t \ln (13 \unit{Gyr}/10\unit{Myr})} & t \in [10\unit{Myr},13\unit{Gyr}]
\end{cases} 
\end{eqnarray}
\abbrvPSgrb{} and \abbrvPSellipticals{} show this distribution is reasonable approximations to compact binary delay
time distributions.\footnote{Simulations suggest the delay time distribution varies from model to model and with mass.
  These variations have less impact on our results than the evolving metallicity distribution of star forming gas.}
%
For suitable choices of parameter, our phenomenological response function is qualitatively consistent with the results of detailed simulations of binary
evolution  \citep{2010ApJ...715L.138B,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,popsyn-LowMetallicityImpact2-StarTrackRevised-2012}.

For binary black holes formed near solar metallicity, the delay time distribution can favor long delays between birth
and merger, as demonstrated in \abbrvPSellipticals{}.

Figure 2 of  \cite{2016arXiv160508783L} is an extreme example of this well-known trend

\editremark{Note}:

\section{Formation rate for binary black holes}
\label{sec:results:BBH}

@@

* BBH properties far from apparent.  No surprise: delays can be long \cite{PSellipticals,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,gwastro-EventPopsynPaper-2016}, particularly because formation in low Z favored,
so system changes significantly from assembly conditions

* EM counterpart properties: delay times are shorter, so more of a surprise that formation scenario matters.  However,
BH properties depend on $Z$, which can influence the mass and spin of BHs that produce EM counterparts.

* Detailed analysis of one host galaxy presented in this work is complementary to the population-based approach used in
\cite{2016arXiv160508783L}, which favored formation in a wide range of galaxies and event times.   % that analysis also
                                % used a poor input from BSE, about the coalescence delay

\appendix
